\bf Adapting the Chumbley Score to match striae on Land Engraved Areas of bullets

### 2018-08-17 doc/paper.Rmd
source of data in table

\begin{table}[ht]
\centering
\resizebox{\textwidth}{!}{\begin{tabular}{rlrrr}
  \hline
Method & Data Source & False Positives & False Negatives\\
  \hline
  \textbf{\citet{aoas}} & LEAs \\
Consecutive Matching Bullets Striae (CMS) &  Ruger &  6.25\% &  33.85\% \\ 
Consecutive Non-matching Striae (CNMS)  & Hamby 252 &  6.26\% & 35.42\%  \\ 
Average Distance (D) & Winchester  &   6.25\% & 45.83\% \\ 
Cross-correlation Function (CCF) & &  6.25\% & 17.71\%  \\ 
Sum of Peaks (S) &  &  6.25\% &  18.23\% \\
   \hline
   \textbf{\citet{chu2013}} & LEAs \\
Consecutive Matching Bullets Striae (CMS) & Ruger &  0\% &  52\% \\ 
   \hline
   \textbf{\citet{chu2010}} \& \textbf{\citet{ma2004}} & LEAs \\
Cross-Correlation Function (CCF) & Ruger  &  - &  - \\ 
    \hline
\end{tabular}}
\caption{\label{tab:bullets-ER} Error Rates in same-source single feature land-to-land analysis reported in the literature. \gk{Note that differences in error rates do not only come from different methods but also differently marking firearms and types of ammunition}}
\end{table}

>>>>

\begin{table}[ht]
\centering
\resizebox{\textwidth}{!}{\begin{tabular}{rlrrr}
  \hline
Method & Data Source & False Positives & False Negatives\\
  \hline
  \textbf{\citet{aoas}} & Hamby 252 (NIST scans) \\
Consecutive Matching Bullets Striae (CMS) &   &  6.25\% &  33.85\% \\ 
Consecutive Non-matching Striae (CNMS)  &  &  6.26\% & 35.42\%  \\ 
Average Distance (D) & Winchester  &   6.25\% & 45.83\% \\ 
Cross-correlation Function (CCF) & &  6.25\% & 17.71\%  \\ 
Sum of Peaks (S) &  &  6.25\% &  18.23\% \\
   \hline
   \textbf{\citet{chu2013}} & Hamby 252 (NIST scans) \\
Consecutive Matching Bullets Striae (CMS) &  &  0\% &  52\% \\ 
   \hline
   \textbf{\citet{chu2010}} & \small{6 types of firearms, 2 types of ammunition} \\
   \textbf{\citet{ma2004}} & \small{NIST standard bullet SRM 8240} \\
Cross-Correlation Function (CCF) &   &  - &  - \\ 
    \hline
\end{tabular}}
\caption{\label{tab:bullets-ER} Error Rates in same-source single feature land-to-land analysis reported in the literature. \gk{Note that differences in error rates do not only come from different methods but also differently marking firearms and types of ammunition}}
\end{table}

### 2018-08-17 doc/paper.tex
reference adher(?) to style guide
adding modified bst file
update reference to aoas paper
source of data in table

% !TeX program = pdfLaTeX
\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{graphicx,psfrag,epsf}
\usepackage{enumerate}
\usepackage{natbib}
\setcitestyle{numbers}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{adjustbox}
\usepackage{multirow}
\usepackage{subfig}
\usepackage[table]{xcolor}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{hyperref}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}

\usepackage{url} % not crucial - just used below for the URL

@@

\tableofcontents
\newpage

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Same-source analyses are a major part of a Forensic Toolmark Examiner's
job. In current practice, examiners make these comparisons by visual
inspection under a comparison microscope and come to one of the
following four conclusions: identification, inconclusive, elimination or
unsuitable for examination \cite{afte-toolmarks1998}. These conclusions
are made on the basis of ``unique surface contours'' of the two
toolmarks being in ``sufficient agreement'' \cite{afte-toolmarks1998}.
AFTE describes the term ``sufficient agreement'' as the possibility of
another tool producing the markings under comparison, as practically
impossible \cite{afte-toolmarks1998}. Potential subject bias in the
assessment as well as the lack of specified error rates are the main
points of criticisms first raised by the National Research Council in
2009 \cite{NAS:2009} and later emphasized further by the President's
Council of Advisors on Science and Technology \cite{pcast2016}.

Technological advances, such as profilometers and confocal microscopy
allow to measure 3D surfaces in a high-resolution digitized form. This
technology has become more accessible over the last decade, and has made
its way into topological images of ballistics evidence, such as bullet
lands and breech faces
\cite{DeKinder1, DeKinder2, Bachrach1, vorburger2016}. Digitized images
of 3D surfaces of form the data basis of statistical analysis of
toolmarks. A statistical approach based on data removes both
subjectivity from the assessment and allows a quantification of error
rates for both false positive and false negative identifications.

@@

Methods for matching marks for a variety of tools have been studied in
the literature (see \autoref{tab:toolmarks-ER} for an overview): in
\cite{manytoolmarks1} and \cite{chumbley} digitized screwdriver marks
have been analyzed using a profilometer; in \cite{manytoolmarks2} 3D
marks from screwdriver, tongue and groove pliers captured using a
confocal microscope have been investigated; digitized marks from
slip-joint pliers generated by a surface profilometer have been
investigated in \cite{afte-chumbley}.

Analysis of these digitized markings require the use of statistical
methods which can quantify the scientific mechanism of comparing
markings and serve as basis for an error rate calculation. In
\cite{manytoolmarks2} a relative distance metric is defined and used as
similarity measure between two toolmarks. This approach is expanded in
\citet{manytoolmarks1}: a set of small segments in the markings of two
toolmarks are extracted and similarity is compared using a maximum
Pearson correlation coefficient. The Chumbley scoring method, first
introduced in \cite{chumbley}, uses a similar but more extensive
framework based on a Mann-Whitney U test of the resulting correlation
coefficients. This approach is non-deterministic, because segments are
chosen randomly. In \cite{hadler} the score is fixed to be deterministic
for each pair of toolmarks by choosing segments for comparison
systematically. This approach also ensures independence between segments
of striae.

\begin{table}[ht]
\centering
\resizebox{\textwidth}{!}{\begin{tabular}{rlrrr}
  \hline
Research paper & Method & Data Source & False Positives & False Negatives\\
  \hline
 \textbf{Faden et al. (2007)} \cite{manytoolmarks1} & Maximum Pearson & Screwdrivers & & \\ & Correlation &  & - & - \\ \hline
 \textbf{Chumbley et al. (2010)} \cite{chumbley} & Randomized & Screwdrivers & & \\ (Same-Surface Same-Angle) & Chumbley Score & & 2.3\% & 8.9\%  \\ \hline
  \textbf{Grieve et al. (2014)}
 \cite{afte-chumbley} & Randomized & Slip-joint  &  & \\ & Chumbley Score &  & - & - \\ \hline 
  \textbf{Hadler \& Morris (2017)} \cite{hadler} & Deterministic & Screwdrivers & & \\ (Same-Surface Same-Angle) & Chumbley Score & & 0\% & 6\%  \\ \hline
  \textbf{Bachrach et al. (2010)} \cite{manytoolmarks2} & Similarity Measure & Screwdrivers & & \\ (Different Surfaces-same angle) & Relative Distance Metric & & 5.9\% &  9.4\% \\ (Same Surfaces-same angle)&  & & 0.22\% &  0\% \\
   \hline
\end{tabular}}
\caption{\label{tab:toolmarks-ER} Error Rates in same-source toolmark analysis reported in the literature. The top four reported papers use some variation of the Chumbley-score method.}
\end{table}

@@

\begin{table}[ht]
\centering
\resizebox{\textwidth}{!}{\begin{tabular}{rlrrr}
  \hline
Method & Data Source & False Positives & False Negatives\\
  \hline
  \textbf{\citet{aoas}} & LEAs \\
Consecutive Matching Bullets Striae (CMS) &  &  6.25\% &  33.85\% \\ 
Consecutive Non-matching Striae (CNMS)  & &  6.26\% & 35.42\%  \\ 
Average Distance (D) &   &   6.25\% & 45.83\% \\ 
Cross-correlation Function (CCF) & &  6.25\% & 17.71\%  \\ 
Sum of Peaks (S) &  &  6.25\% &  18.23\% \\
   \hline
   \textbf{\citet{chu2013}} & LEAs \\
Consecutive Matching Bullets Striae (CMS) &  &  0\% &  52\% \\ 
   \hline
   \textbf{\citet{chu2010}} \& \textbf{\citet{ma2004}} & LEAs \\
Cross-Correlation Function (CCF) &  &  - &  - \\ 
    \hline
\end{tabular}}
\caption{\label{tab:bullets-ER} Error Rates in same-source single feature land-to-land analysis reported in the literature.}
\end{table}

@@

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{scans-for-land-engraved-areas}{%
\subsection{Scans for land engraved
areas}\label{scans-for-land-engraved-areas}}

Comparisons of striae from bullets are usually based on comparisons of
striae in land engraved areas, which are extracted in form of cross
sections, called \emph{profiles} \citep{aoas,ma2004}. From profiles
bullet \emph{signatures} \citep{chu2013,aoas} are extracted as residuals
of a loess fit or Gaussian filter. This effectively removes topographic
structure from the data in the attempt to increase the signal to noise
ratio. \autoref{fig:rgl} shows how the signature from a bullet land
(bottom) lines up with the image of the land (top) from which it was
extracted. We can see in the figure how the depth and relative position
of the striation markings seen in the image are interpreted as peaks and
valleys in the signature.

@@

\hypertarget{the-chumbley-score-test}{%
\subsection{The Chumbley Score Test}\label{the-chumbley-score-test}}

@@

\hypertarget{a-problem-with-failed-tests}{%
\subsection{A problem with failed
tests}\label{a-problem-with-failed-tests}}

@@

\hypertarget{a-modified-approach}{%
\subsection{A modified approach}\label{a-modified-approach}}

@@

\hypertarget{testing-setup}{%
\subsection{Testing setup}\label{testing-setup}}

\hypertarget{the-data}{%
\subsubsection{The Data}\label{the-data}}

@@

\hypertarget{setup}{%
\subsubsection{Setup}\label{setup}}

@@

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{failed-tests}{%
\subsection{Failed Tests}\label{failed-tests}}

@@

\hypertarget{coarseness}{%
\subsection{Coarseness}\label{coarseness}}

@@

\hypertarget{error-rate-assessment}{%
\subsection{Error rate assessment}\label{error-rate-assessment}}

@@

\hypertarget{observed-versus-nominal-type-i-error-rates}{%
\subsection{Observed versus Nominal Type I error
rates}\label{observed-versus-nominal-type-i-error-rates}}

@@

\hypertarget{high-resolution-hamby-44-scans}{%
\subsection{High resolution Hamby 44
scans}\label{high-resolution-hamby-44-scans}}

@@

\hypertarget{conclusions}{%
\section{Conclusions}\label{conclusions}}

@@

\hypertarget{appendix}{%
\section{Appendix}\label{appendix}}

@@

\bibliographystyle{vancouver}
\bibliography{bibliography}

>>>>

% !TeX program = pdfLaTeX
\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{graphicx,psfrag,epsf}
\usepackage{enumerate}
%\usepackage[numbers]{natbib}
\usepackage{natbib}
\setcitestyle{numbers} 
\makeatletter % Reference list option change
\renewcommand\@biblabel[1]{#1. } % from [1] to 1
\makeatother %

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{adjustbox}
\usepackage{multirow}
\usepackage{subfig}
\usepackage[table,xcdraw]{xcolor}
\usepackage{wrapfig}
\usepackage{float}
%\usepackage{colortbl}
\usepackage[colorlinks]{hyperref}
\hypersetup{
  colorlinks=true,
  citecolor=black,
  linkcolor=black,
  urlcolor=blue}

\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{url} % not crucial - just used below for the URL

\usepackage{etoolbox}% http://ctan.org/pkg/etoolbox
\makeatletter
\patchcmd{\subsection}{\bfseries}{\relax}{}{}% Non-bold \subsection
\patchcmd{\subsubsection}{\bfseries}{\relax}{}{}% Non-bold \subsection
\makeatother

@@

\newpage

Same-source analyses are a major part of a Forensic Toolmark Examiner's
job. In current practice, examiners make these comparisons by visual
inspection under a comparison microscope and come to one of the
following four conclusions: identification, inconclusive, elimination or
unsuitable for examination \cite{afte-toolmarks1998}. These conclusions
are made on the basis of ``unique surface contours'' of the two
toolmarks being in ``sufficient agreement'' \cite{afte-toolmarks1998}.
AFTE describes the term ``sufficient agreement'' as the possibility of
another tool producing the markings under comparison, as practically
impossible \cite{afte-toolmarks1998}. Potential subject bias in the
assessment as well as the lack of specified error rates are the main
points of criticisms first raised by the National Research Council in
2009 \cite{NAS:2009} and later emphasized further by the President's
Council of Advisors on Science and Technology \cite{pcast2016}.
{\textcolor{blue}{need a sentence on striated marks}}

Technological advances, such as profilometers and confocal microscopy
allow {\textcolor{blue}{one}} to measure 3D surfaces in a
high-resolution digitized form. This technology has become more
accessible over the last decade, and has made its way into topological
images of ballistics evidence, such as bullet lands and breech faces
\cite{DeKinder1, DeKinder2, Bachrach1, vorburger2016}.
{\textcolor{blue}{small change}}Digitized images of 3D surfaces, form
the basis of statistical analysis of toolmarks. A statistical approach
based on data removes both subjectivity from the assessment and allows a
quantification of error rates for both false positive and false negative
identifications.

@@

Methods for matching {\textcolor{blue}{striated}} marks for a variety of
tools have been studied in the literature (see
\autoref{tab:toolmarks-ER} for an overview): in \cite{manytoolmarks1}
and \cite{chumbley} digitized screwdriver marks have been analyzed using
a profilometer; in \cite{manytoolmarks2} 3D marks from screwdriver,
tongue and groove pliers captured using a confocal microscope have been
investigated; digitized marks from slip-joint pliers generated by a
surface profilometer have been investigated in \cite{afte-chumbley}.

{\textcolor{blue}{changed}} These digitized markings allow the use of
statistical methods which can quantify the scientific mechanism of
comparing markings and serve as basis for an error rate calculation. In
\cite{manytoolmarks2} a relative distance metric is defined and used as
similarity measure between two toolmarks. This approach is expanded in
\citet{manytoolmarks1}: a set of small segments in the markings of two
toolmarks are extracted and similarity is compared using a maximum
Pearson correlation coefficient. The Chumbley scoring method, first
introduced in \cite{chumbley}, uses a similar but more extensive
framework based on a Mann-Whitney U test of the resulting correlation
coefficients. This approach is non-deterministic, because segments are
chosen randomly. In \cite{hadler} the score is fixed to be deterministic
for each pair of toolmarks by choosing segments for comparison
systematically. This approach also ensures independence between segments
of striae.

\begin{table}[ht]
\centering
\resizebox{\textwidth}{!}{\begin{tabular}{rlrrr}
  \hline
Research paper & Method & Data Source & False Positives & False Negatives\\
  \hline
 \textbf{Faden et al. (2007)} \cite{manytoolmarks1} & Maximum Pearson & Screwdrivers & & \\ & Correlation &  & - & - \\ \hline
 \textbf{Chumbley et al. (2010)} \cite{chumbley} & Randomized & Screwdrivers & & \\ (Same-Surface Same-Angle) & Chumbley Score & & 2.3\% & 8.9\%  \\ \hline
  \textbf{Grieve et al. (2014)}
 \cite{afte-chumbley} & Randomized & Slip-joint  &  & \\ & Chumbley Score &  & - & - \\ \hline 
  \textbf{Hadler \& Morris (2017)} \cite{hadler} & Deterministic & Screwdrivers & & \\ (Same-Surface Same-Angle) & Chumbley Score & & 0\% & 6\%  \\ \hline
  \textbf{Bachrach et al. (2010)} \cite{manytoolmarks2} & Similarity Measure & Screwdrivers & & \\ (Different Surfaces-same angle) & Relative Distance Metric & & 5.9\% &  9.4\% \\ (Same Surfaces-same angle)&  & & 0.22\% &  0\% \\
   \hline
\end{tabular}}
\caption{\label{tab:toolmarks-ER} Error Rates in same-source {\textcolor{blue}{striated}} toolmark analysis reported in the literature. The top four reported papers use some variation of the Chumbley-score method.}
\end{table}

@@

\begin{table}[ht]
\centering
\resizebox{\textwidth}{!}{\begin{tabular}{rlrrr}
  \hline
Method & Data Source & False Positives & False Negatives\\
  \hline
  \textbf{\citet{aoas}} & Hamby 252 (NIST scans) \\
Consecutive Matching Bullets Striae (CMS) &   &  6.25\% &  33.85\% \\ 
Consecutive Non-matching Striae (CNMS)  &  &  6.26\% & 35.42\%  \\ 
Average Distance (D) & Winchester  &   6.25\% & 45.83\% \\ 
Cross-correlation Function (CCF) & &  6.25\% & 17.71\%  \\ 
Sum of Peaks (S) &  &  6.25\% &  18.23\% \\
   \hline
   \textbf{\citet{chu2013}} & Hamby 252 (NIST scans) \\
Consecutive Matching Bullets Striae (CMS) &  &  0\% &  52\% \\ 
   \hline
   \textbf{\citet{chu2010}} & \small{6 types of firearms, 2 types of ammunition} \\
   \textbf{\citet{ma2004}} & \small{NIST standard bullet SRM 8240} \\
Cross-Correlation Function (CCF) &   &  - &  - \\ 
    \hline
\end{tabular}}
\caption{\label{tab:bullets-ER} Error Rates in same-source single feature land-to-land analysis reported in the literature. {\textcolor{blue}{Note that differences in error rates do not only come from different methods but also differently marking firearms and types of ammunition}}}
\end{table}

@@

\hypertarget{methods}{%
\section*{Methods}\label{methods}}
\addcontentsline{toc}{section}{Methods}

\hypertarget{scans-for-land-engraved-areas}{%
\subsection*{\texorpdfstring{\emph{Scans for land engraved
areas}}{Scans for land engraved areas}}\label{scans-for-land-engraved-areas}}
\addcontentsline{toc}{subsection}{\emph{Scans for land engraved areas}}

{\textcolor{blue}{Bullet striae are most pronounced at the bottom of the bullet heel (because the heel has the most contact with the insides of the barrel). Howver, these areas are also affected by the break off due to friction effects between barrel and the bullet. An optimal crosscut is chosen orthogonally to the striae, close to the heel while avoiding break off, see also \citep{aoas}}}.
Comparisons of striae from bullets are usually based on comparisons of
striae in land engraved areas, which are extracted in form of cross
sections, called \emph{profiles} \citep{aoas,ma2004}. From profiles
bullet \emph{signatures} \citep{chu2013,aoas} are extracted as residuals
of a loess fit or Gaussian filter. This effectively removes topographic
structure from the data in the attempt to increase the signal to noise
ratio. \autoref{fig:rgl} shows how the signature from a bullet land
(bottom) lines up with the image of the land (top) from which it was
extracted. We can see in the figure how the depth and relative position
of the striation markings seen in the image are interpreted as peaks and
valleys in the signature.

@@

\hypertarget{the-chumbley-score-test}{%
\subsection*{\texorpdfstring{\emph{The Chumbley Score
Test}}{The Chumbley Score Test}}\label{the-chumbley-score-test}}
\addcontentsline{toc}{subsection}{\emph{The Chumbley Score Test}}

@@

\hypertarget{a-problem-with-failed-tests}{%
\subsection*{\texorpdfstring{\emph{A problem with failed
tests}}{A problem with failed tests}}\label{a-problem-with-failed-tests}}
\addcontentsline{toc}{subsection}{\emph{A problem with failed tests}}

@@

\hypertarget{a-modified-approach}{%
\subsection*{\texorpdfstring{\emph{A modified
approach}}{A modified approach}}\label{a-modified-approach}}
\addcontentsline{toc}{subsection}{\emph{A modified approach}}

@@

\hypertarget{testing-setup}{%
\subsection*{\texorpdfstring{\emph{Testing
setup}}{Testing setup}}\label{testing-setup}}
\addcontentsline{toc}{subsection}{\emph{Testing setup}}

\hypertarget{the-data}{%
\subsubsection*{The Data}\label{the-data}}
\addcontentsline{toc}{subsubsection}{The Data}

@@

\hypertarget{setup}{%
\subsubsection*{Setup}\label{setup}}
\addcontentsline{toc}{subsubsection}{Setup}

@@

\hypertarget{results}{%
\section*{Results}\label{results}}
\addcontentsline{toc}{section}{Results}

\hypertarget{failed-tests}{%
\subsection*{\texorpdfstring{\emph{Failed
Tests}}{Failed Tests}}\label{failed-tests}}
\addcontentsline{toc}{subsection}{\emph{Failed Tests}}

@@

\hypertarget{coarseness}{%
\subsection*{\texorpdfstring{\emph{Coarseness}}{Coarseness}}\label{coarseness}}
\addcontentsline{toc}{subsection}{\emph{Coarseness}}

@@

\hypertarget{error-rate-assessment}{%
\subsection*{\texorpdfstring{\emph{Error rate
assessment}}{Error rate assessment}}\label{error-rate-assessment}}
\addcontentsline{toc}{subsection}{\emph{Error rate assessment}}

@@

\hypertarget{observed-versus-nominal-type-i-error-rates}{%
\subsection*{\texorpdfstring{\emph{Observed versus Nominal Type I error
rates}}{Observed versus Nominal Type I error rates}}\label{observed-versus-nominal-type-i-error-rates}}
\addcontentsline{toc}{subsection}{\emph{Observed versus Nominal Type I
error rates}}

@@

\hypertarget{high-resolution-hamby-44-scans}{%
\subsection*{\texorpdfstring{\emph{High resolution Hamby 44
scans}}{High resolution Hamby 44 scans}}\label{high-resolution-hamby-44-scans}}
\addcontentsline{toc}{subsection}{\emph{High resolution Hamby 44 scans}}

@@

\hypertarget{conclusions}{%
\section*{Conclusions}\label{conclusions}}
\addcontentsline{toc}{section}{Conclusions}

@@

\hypertarget{appendix}{%
\section*{Appendix}\label{appendix}}
\addcontentsline{toc}{section}{Appendix}

@@

\bibliographystyle{jfs-authoryear}
\bibliography{bibliography}
